---
title: "Assignment 2"
subtitle: |
  Data Acquisition and Survey Methods  
  Survey Analysis Report
author: |
  Group 39
date: "2026"
output:
  pdf_document:
    toc: true
    number_sections: true
---

------------------------------------------------------------------------

```{r setup, include=FALSE}
library(tidyverse)
library(janitor)
library(knitr)

knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE
)
```

# Introduction

This report analyzes the relationship between meal habits and perceived study productivity among university students. The focus is on three research questions related to breakfast frequency, the number of main meals per day, and self-reported study productivity.

The analysis focuses on the following three research questions and hypotheses:

-   **RQ1:** Is the frequency of eating breakfast associated with students' perceived study productivity?\
    **H1:** Students who eat breakfast more often report higher perceived study productivity.

-   **RQ2:** Is the number of main meals per day associated with students' perceived study productivity?\
    **H2:** Students who eat more main meals per day report higher perceived study productivity.

-   **RQ3:** Is the frequency of eating breakfast associated with the number of main meals students usually eat per day?\
    **H3:** Students who eat breakfast more often also report eating more main meals per day.

For the analysis, only the variables relevant to Group 39 were selected from the full survey dataset.

# Data Preparation

```{r}
# Load the survey data and keep the original column names
data <- read.csv("../data/Survey_results_2026.csv", check.names = FALSE)

# Select only the columns relevant for our group
group39_data <- data[, c(
  "DQ How old are you?",
  "DQ What is your gender?",
  "DQ What is your academic program?",
  "G39 How often do you eat breakfast during a typical study week?",
  "G39 How many main meals do you usually eat per day, excluding snacks?",
  "G39 How productive do you usually feel during your study sessions?"
)]

# Rename the columns to shorter names
names(group39_data) <- c(
  "age",
  "gender",
  "academic_program",
  "breakfast_frequency",
  "main_meals_per_day",
  "study_productivity"
)

# Convert breakfast frequency to an ordered factor
group39_data <- group39_data %>%
  mutate(
    main_meals_per_day = as.numeric(main_meals_per_day),
    study_productivity = as.numeric(study_productivity),
    breakfast_frequency = factor(
      breakfast_frequency,
      levels = c(
        "Never",
        "1-2 days per week",
        "3-4 days per week",
        "5-6 days per week",
        "Every day"
      ),
      ordered = TRUE
    ),
    breakfast_score = as.numeric(breakfast_frequency) - 1
  )

# Show the first rows
head(group39_data)
```

The selected dataset contains demographic variables as well as the three variables needed to test the research hypotheses. The variables were converted into appropriate formats for the analysis. The number of main meals per day and study productivity were transformed into numeric variables. Breakfast frequency was converted into an ordered factor because the response categories follow a natural order from "Never" to "Every day". Additionally, a numeric breakfast score was created to use breakfast frequency in the statistical correlation tests.

```{r}
str(group39_data)
colSums(is.na(group39_data))
```

The dataset contains 209 observations and 7 variables after data preparation. No missing values were found in any of the selected variables, so no observations had to be removed before the analysis.

# Hypothesis 1: Breakfast Frequency and Study Productivity

**Research Question:** Is the frequency of eating breakfast associated with students' perceived study productivity?
**Hypothesis:** Students who eat breakfast more often report higher perceived study productivity.

## Exploratory Data Analysis:

Choose appropriate methods to explore the categorical and quantitative features in your data set (one visualization per research question). Briefly explain the methods and comment on your findings.

```{r}
# Calculate average study productivity for each breakfast category
h1_summary <- aggregate(
  study_productivity ~ breakfast_frequency,
  data = group39_data,
  FUN = mean
)

# Show the summary table
h1_summary
```

```{r}
# Create a barplot for Hypothesis 1
barplot(
  h1_summary$study_productivity,
  names.arg = h1_summary$breakfast_frequency,
  xlab = "Breakfast frequency",
  ylab = "Average study productivity",
  main = "Average study productivity by breakfast frequency",
  las = 2
)
```

## Descriptive Inference:

Compute suitable summary statistics and display the results in a table.

## Analytic Inference:

Select an appropriate test statistic to test your hypothesis. Summarize the results and explain how your findings are connected to your research question.

# Hypothesis 2: Main Meals per Day and Study Productivity

## Exploratory Data Analysis:

## Descriptive Inference:

## Analytic Inference:

# Hypothesis 3: Breakfast Frequency and Main Meals per Day

## Exploratory Data Analysis:

## Descriptive Inference:

## Analytic Inference:

# Explorative Analsis

Mediator Hyotheses

# Summary of Findings

# Conclusion
