---
title: "Assignment 2"
subtitle: |
  Data Acquisition and Survey Methods  
  Survey Analysis Report
author: |
  Group 39
date: "2026"
output:
  pdf_document:
    toc: true
    number_sections: true
---

------------------------------------------------------------------------

```{r, include=FALSE}
library(tidyverse)
library(janitor)
library(knitr)
```

# Introduction

This report analyzes the relationship between meal habits and perceived study productivity among university students. The focus is on three research questions related to breakfast frequency, the number of main meals per day, and self-reported study productivity.

For the analysis, only the variables relevant to Group 39 were selected from the full survey dataset.

# Data Preparation

```{r}
# Load the survey data and keep the original column names
data <- read.csv("../data/Survey_results_2026.csv", check.names = FALSE)

# Select only the columns relevant for our group
group39_data <- data[, c(
  "DQ How old are you?",
  "DQ What is your gender?",
  "DQ What is your academic program?",
  "G39 How often do you eat breakfast during a typical study week?",
  "G39 How many main meals do you usually eat per day, excluding snacks?",
  "G39 How productive do you usually feel during your study sessions?"
)]

# Rename the columns to shorter names
names(group39_data) <- c(
  "age",
  "gender",
  "academic_program",
  "breakfast_frequency",
  "main_meals_per_day",
  "study_productivity"
)

# Show the first rows
head(group39_data)
```

The selected dataset contains demographic variables as well as the three variables needed to test the research hypotheses.

# Hypothesis 1: Breakfast Frequency and Study Productivity

Research Question: Is the frequency of eating breakfast associated with students' perceived study productivity?

Hypothesis: Students who eat breakfast more often report higher perceived study productivity.

```{r}
# Make sure study productivity is numeric
group39_data$study_productivity <- as.numeric(group39_data$study_productivity)

# Define the correct order for breakfast frequency
group39_data$breakfast_frequency <- factor(
  group39_data$breakfast_frequency,
  levels = c(
    "Never",
    "1-2 days per week",
    "3-4 days per week",
    "5-6 days per week",
    "Every day"
  ),
  ordered = TRUE
)

# Calculate average study productivity for each breakfast category
h1_summary <- aggregate(
  study_productivity ~ breakfast_frequency,
  data = group39_data,
  FUN = mean
)

# Show the summary table
h1_summary
```

```{r}
# Create a barplot for Hypothesis 1
barplot(
  h1_summary$study_productivity,
  names.arg = h1_summary$breakfast_frequency,
  xlab = "Breakfast frequency",
  ylab = "Average study productivity",
  main = "Average study productivity by breakfast frequency",
  las = 2
)
```


# Hypothesis 2: Main Meals per Day and Study Productivity


# Hypothesis 3: Breakfast Frequency and Main Meals per Day


# Summary of Findings

# Conclusion
